---
title: "Topic 4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Topic 4- Regression to the mean
```{r}
library(SportsAnalytics270)
data("nba_misc")
head(nba_misc)
```

```{r}
nba_misc$WR<-nba_misc$W/82

var(nba_misc$WR)
```
```{r}
0.5*0.5/82
```
```{r}
data(nba_east)
data(nba_west)

nba_st<- rbind(nba_east, nba_west)



nba_st$PW <- nba_st$PS^14.105/(nba_st$PS^14.105+nba_st$PA^14.105)
nba_st$G <- nba_st$W+nba_st$L
nba_st$PW_G <- round(nba_st$PW*nba_st$G)


nba_st$DF <- nba_st$PW_G-nba_st$W

#pthagorea wins- wins
```

```{r}
library(ggplot2)
library(dplyr)
nba_st %>% filter(Team=="Sacramento Kings") %>%
  ggplot(aes(x=Season, y=DF))+geom_point()+geom_line()+
  ggtitle('Wins difference for Sacramento Kings (Luck)')
```

```{r}
means <- nba_st %>% group_by(Team) %>% summarise(mean=mean(DF))
head(means)
```
```{r}
ggplot(means, aes(x=mean))+geom_histogram()
```
#Competitive Balance- how competitive are leagues?
```{r}
data("seriea_st")
colnames(seriea_st)
head(seriea_st)
```

```{r}
seriea_s <- seriea_st %>%
  group_by(Season) %>%
  summarise(SD=sd(Pt))
seriea_s
```

```{r}
ggplot(seriea_s, aes(Season, SD))+geom_bar(stat="identity")+
  ggtitle("SD of Serie A by season")+
  coord_cartesian(ylim = c(10,22))
```

```{r}
data(nfl_st)
summary(nfl_st)
```
```{r}
library(dplyr)
nfl_s<-nfl_st %>%
  group_by(Season) %>%
  summarise(SD=sd(Wpct))
nfl_s
```
```{r}
ggplot(nfl_s, aes(Season, SD))+geom_bar(stat="identity")+
  ggtitle("NFL SD by season")+
  coord_cartesian(ylim = c(0.1,0.25))
```
```{r}
#Ideal StDev
id_s <- 0.5/sqrt(16)

#win pct stdev
sd_wpct<-sd(nfl_st$Wpct)

#noll-scully number for whole data set
sd_wpct/id_s
```

```{r}
library(dplyr)

nfl_s<-nfl_st %>%
  group_by(Season) %>%
  summarise(NS=sd(Wpct)/id_s)
nfl_s
```

```{r}
library(ggplot2)
ggplot(nfl_s, aes(Season, NS))+geom_bar(stat="identity")+
  coord_cartesian(ylim=c(1,1.75))+
  ggtitle("Noll-Skully number by season")+
  geom_hline(yintercept = mean(nfl_s$NS), col="red", size=1.5)
```
```{r}
#Concentration indexes- concentarted among teams in league, (points earned by top clubs/ total number points earned for all clubs)

#increasing means competitive is decreasing

```

```{r}
top5<- seriea_st %>%
  group_by(Season) %>%
  filter(Rank<6) %>%
  summarise(TopP=sum(Pt))

all<- seriea_st %>%
      group_by(Season) %>%
  summarize(Pt=sum(Pt))

C5<-data.frame(Season=top5$Season, C5=top5$topP/all$Pt)
```

```{r}
ggplot(C5, aes(Season, C5))+geom_bar(stat="identity")+
  ggtitle("C% index for Serie A 1996-2018")+
  coord_cartesian(ylim=c(0.2,0.42))
```
```{r}
ggplot(C5, aes(Season,1- C5))+geom_bar(stat="identity")+
  ggtitle("C% index for Serie A 1996-2018")+
  coord_cartesian(ylim=c(0.4,0.65))
```
```{r}
champ_c<- seriea_st %>%
  filter(Rank == 1) %>%
  group_by(Team) %>%
  summarize(champ=n()) %>%
  arrange(desc(champ))

champ_c
```

```{r}
#CI4 top 4 teams for championships of all last 25 years, if CI5 not less than 100%

100*sum(champ_c[1:4,2])/sum(champ_c$champ)
```
```{r}
#HHI- sahre of each team- concentration

HHI<-seriea_st %>%
  group_by(Season) %>%
  mutate(Perc=Pt/sum(Pt)) %>%
  summarise(HHI = sum(Perc^2))
```

```{r}
ggplot(HHI, aes(Season, HHI))+geom_bar(stat="identity")+
  ggtitle("HHI index by season")+
  coord_cartesian(ylim=c(0,0.065))
```

```{r}
seriea_st %>%
  group_by(Season) %>%
  summarise(count=n())
```
#which to take? correlate with other variable
```{r}
dHHI<-seriea_st %>%
  group_by(Season) %>%
  mutate(Perc=Pt/sum(Pt)) %>%
  summarise(HHI = sum(Perc^2)-1/n())
```
```{r}
ggplot(dHHI, aes(Season, HHI))+geom_bar(stat="identity")+
  ggtitle("dHHI index by season")
```
#low value means competition
```{r}
library(SportsAnalytics270)
str(mlb_standings)
```

```{r}
mlb_HHI <- mlb_standings %>%
  group_by(Season) %>%
  summarise(HHI=sum(W^2)*(4/(30^2*162^2)))
```
```{r}
ggplot(mlb_HHI, aes(Season, HHI))+geom_bar(stat="identity")+
  geom_hline(yintercept = 1/30, col="red", size=1.5)+
  coord_cartesian(ylim = c(0.03,0.035))+
  ggtitle("HHI index by season for mlb")
```

#competitive is more or less the same
#leading indicator-cause, laggingni-because of

```{r}
library(dplyr)
library(SportsAnalytics270)
data("mlb_salary")
str(mlb_salary)
```
```{r}
mlb_s<-mlb_salary %>%
  group_by(Season) %>%
  summarise(Tot.salary=sum(Current), SD=sd(Current))

cor(mlb_s$Tot.salary, mlb_HHI$HHI)
cor(mlb_s$SD, mlb_HHI$HHI)
```
```{r}
mlb_HHI_s<-mlb_salary %>%
  group_by(Season) %>%
  mutate(Perc=Current/sum(Current)) %>%
  summarise(HHI_S=sum(Perc^2))
```
```{r}
cor(mlb_HHI_s$HHI_S, mlb_HHI$HHI)
```

```{r}
#Short run competitiveness
```

```{r}
data("f_data_sm")
f_2009 <- f_data_sm %>%
  filter(SEASON > 2008,
         COUNTRY %in% c("Italy", "Germany", "Spain", "England",
                        "Netherlands", "Portugal"), !is.na(H))
head(f_2009)
```
prob of HT win 1/1.2,  away prob 1/15
above one is marginal book maker
```{r}
install.packages("odds.converter")
library(odds.converter)
x <- odds.fv(c(1.4,3.8,8), input = 'dec', output = 'prob')
x
```
above is odds free of vigorish
```{r}
sum(x)
```
```{r}
f_2009[,c("H", "D", "A")] <- odds.fv(f_2009[,c("H", "D", "A")],
                                     input = 'dec', output = 'prob')

head(f_2009)

#dec is decimal odds, and we want outcome of probabilites
```
```{r}
library(measures)
brier_score <- f_2009 %>% group_by(COUNTRY) %>%
  do(SCORE = multiclass.Brier(.[,c("H", "D", "A")], .$FTR))
brier_score <- as.data.frame(brier_score)
brier_score$SCORE <- unlist(brier_score$SCORE)
```
```{r}
library(ggplot2)
ggplot(brier_score, aes(x=COUNTRY, y=SCORE))+geom_bar(stat='identity')+
  coord_cartesian(ylim = c(0.5,0.6))+ggtitle('Match Uncertainty by country')
```
Germany is most competitive(highest uncertainty)
```{r}
brier_score <- f_2009 %>% group_by(COUNTRY, SEASON) %>%
  do(SCORE = multiclass.Brier(.[,c("H", "D", "A")], .$FTR))
brier_score <- as.data.frame(brier_score)
brier_score$SCORE <- unlist(brier_score$SCORE)
```

```{r}
brier_score %>% 
  ggplot(aes(x=SEASON, y=SCORE))+
  geom_bar(stat='identity', position=position_dodge(width=1))+
  facet_grid(~COUNTRY)+coord_flip()+
  scale_x_continuous(breaks = 2009:2019)+
  ggtitle("Brier Score by Country")
```

